---
title: "WDI data project"
author: "Klaas Fiete Krutein"
date: "10/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/Users/HFSM/Documents/University/University of Washington/Courses/2018 Q4/IND E 498/Project")
data <- read.csv("HDIdata_new.csv")
```

At first, we look at our dataset through investigating its structure and summary
```{r}
str(data)
```
As we can see we have a lot of missing values for many of the years. If we want to use a large part of the data set to check for correlations, we need to clean our data set. Thus, we create a function that kills all the variables that do not show data for more than 50% of the time. 
```{r}
library(dplyr)
col_test <- apply(data[,3:length(data)], 2, function(x) sum(is.na(x)))
col_test <- col_test/nrow(data)

data_test <- cbind(data[,1:2], data %>% select(names(col_test[col_test < 0.3])))
head(data_test[5:10,])

#Changes need to be applied here as well
```
Now, we can check how the function works and look at the resulting data frame:
Basic tests on regression modelling:
```{r}
# Create a training data (half the original data size)
train.ix <- sample(nrow(data_test),floor(nrow(data)/2))
data.train <- data_test[train.ix,4:length(data_test)]
GDPno_nas <- rownames(data.train)[!is.na(data.train$NY.GDP.PCAP.KD.ZG)]
data.GDP_nonas <- data.train[GDPno_nas,]

# Create a subset of data with only non-NAs in the response (this is not working properly right now)
col_test2 <- apply(data.GDP_nonas, 2, function(x) sum(is.na(x)))
col_test2 <- col_test2/nrow(data.GDP_nonas)
library(dplyr)
data.train2 <-  data.train %>% select(names(col_test2[col_test2 < 0.2]))

#Changes need to be applied here as welll
# 3:length(data_test)
# Create a testing data (half the original data size)
data.test <- data_test[-train.ix,4:length(data_test)]
```
Build a regression model on this data set
```{r}
lm.AD <- lm(NY.GDP.PCAP.KD.ZG ~ ., data = data.train2)
summary(lm.AD)
# We need to get a data frame that has no pure NA-values for all response related data points.
```


```{r}
lm.AD.reduced <- step(lm.AD, direction="backward", test="F")
```



